---
title: "Untitled"
output: html_document
date: "2024-10-14"
---



```{r}
patient_data = read.table("support2.csv", sep=",", header=TRUE);
patient_data[patient_data == ""] <- NA # certains champs sont vides, remplace par NA
head(patient_data, 500)
```

Les paramètres : 
  age (of patients) - no NA
  death (on 31/12/1994) - no NA - not all patients are followed up because discharged - Target value
  Sex - no NA - values could be binarized
  hospdead - no NA - death in hospital - Target Value
  slos - no NA - (Nb Days from Study Entry to Discharge)
  d.time - no NA - (follow up days)
  dzgroup - no NA -
  dz.class - no NA -
  num.co - no NA -
  edu - years of edu - with NA - not useful
  income - listed values, hard to categorize
  scoma - Support day 3 Coma score based on GSC - with NA
  charges (hospital) - with NA
  totcst - total ratio of costs to charge - not interesting - ~10% of missings values
  totmcst - Total micro cost - not interesting - 38% of missing values
  avtisst - 
  race - 
**  sps - support physiology score on day 3, help track the severity of illness and patient progress over time - 
**  aps - APACH III day 3 physiology score, assess illness severity
**  surv2m - Support model 2-month survival estimate at day 3" refers to the use of a predictive model to estimate a patient's likelihood of surviving for 2 months - Not useful
**  surv6m - same but for 6 months - not useful
  hday - no NA - day at wich patient entered study - not really useful
  diabetes - no NA -
  dementia - no NA -
  ca - cancer - three values possibles (yes,no, metastatic )
**  prg2m - Physician's 2-month survival estimate for patient - not useful
**  prg6m - same for 6 month - not useful
**  dnr - Do not rescuscitate order - not useful
**  dnrday - not useful
  meanbp - mean arterial blood pressure of the patient (day 3)
  wblc - count of white blood cells (day 3) 2%
  hrt - heart rate (day 3)
  temp - temperature measured (day 3)
  pafi - PaO2/FiO2 ratio (Arterial oxygen partial pressure / fractional inspired oxygen)
  alb - serum albumin levels (day 3) , assess liver function, kidney function, nutritional status, and chronic diseases.- 37% of missing values
  bili - bilirubin levels (day 3) , assessing liver function, red blood cell health, and conditions such as jaundice - 28% of missing values
  crea - serum creatinine levels - Monitoring serum creatinine helps diagnose and manage various kidney-related conditions. 25 % of missing values
  sod - serum sodium concentration - reflects the body's fluid balance and is critical for normal physiological functioning. 
  ph - Arterial blood ph - help diagnose lung, kidney, metabolic disease, but can be due to head or neck injuries (affect breathing) 25% of missing values
  glucose - 49% of missing values
  bun - 47% of missing values
  urine - 53% of missing values
  adlp - Activities of Daily Living Higher value More chance of survival -61% of missing values
  adls - Same but filled by a surrogate not the patient - 31% of missing values
  sfdm2 - target
  adlsc - 
  
```{r}
sum(is.na(patient_data$sfdm2)) + sum(is.na(patient_data$race))
```
The diffrents target values are : 
```{r}
Y1 <- patient_data$death
Y2 <- patient_data$hospdead
Y3 <- patient_data$sfdm2
```

a enlever : aps sps sur2m sur6m prg2m prg6m dnr dnrday; à supprimer aussi : edu (pas de relation logique avec le reste), adlp (61% de valeurs manquantes, énorme)

```{r}
colonnes_a_supprimer <- c("aps", "sps", "surv2m", "surv6m", "prg2m", "prg6m", "dnr", "dnrday", "edu", "adlp")
patient_data <- patient_data[, !(names(patient_data) %in% colonnes_a_supprimer)]
head(patient_data)
```
Arranger dzgroup (catégorique), dzclass (catégorique), income (catégorique), race (catégorique), ca
remplir NA :  edu (17.95%), scoma (<1%), charges (2%),  totcst (~10%), totmcst (38%), avtisst (0.9%), meanbp (0.01%), hrt (0.01%), temp (0.01%), sod (0.01%), ph (25.1%), glucose (49%),  adlp (61%), adls (31%)

pafi (25.5%), alb (37%), bili (28.6%), crea (0.7%), bun(47%), wblc (2.3%), urine (53%), income (33%),
comment remplir les valeurs manquantes pour race ? qui est une variable à priori importante

```{r}
# Créer une liste des valeurs normales
standard_values <- list(
  alb = 3.5,    # Serum albumin
  pafi = 333.3, # PaO2/FiO2 ratio
  bili = 1.01,  # Bilirubin
  crea = 1.01,  # Creatinine
  bun = 6.51,   # BUN
  wblc = 9,     # White blood count (en milliers)
  urine = 2502,  # Urine output
  race = "other",
  income = "$11-$25k"  # Salaire moyen aux USA autour de 1990
)

# Fonction pour imputer les valeurs manquantes avec les normales
impute_values <- function(data, colonnes, valeurs) {
  for (col in colonnes) {
    if (col %in% names(data)) {
      data[[col]][is.na(data[[col]])] <- valeurs[[col]]
    }
  }
  return(data)
}

# Imputation par la moyenne pour les colonnes restantes
impute_moy <- function(data) {
  for (col in names(data)) {
    if (any(is.na(data[[col]]))) {  # Vérifie si la colonne contient des NA
      if (is.numeric(data[[col]])) {  # Vérifie si la colonne est numérique
        data[[col]][is.na(data[[col]])] <- mean(data[[col]], na.rm = TRUE)
      }
    }
  }
  return(data)
}


patient_data <- impute_moy(patient_data)

# Liste des colonnes à imputer
columns_to_assign <- names(standard_values)

# Appliquer l'imputation sur votre dataset
patient_data <- impute_values(patient_data, columns_to_assign, standard_values)

# Vérification des valeurs imputées
head(patient_data)

```
  
Encodage des variables catégorique :
  
2. Le sexe de la personne est aussi important, on effet selon qu'on est un homme ou une femme et les effets n'ont pas forcement la même signification et/ou impact.
On a des variables catégoriques, il faut binairiser ces données

```{r}
patient_data$sex <- ifelse(patient_data$sex == "male", 1, 0)

patient_data$race <- as.numeric(factor(patient_data$race, levels = c("missing", "black", "asian", "hispanic", "white", "other")))

patient_data$income <- as.numeric(factor(patient_data$income, levels = c("under $11k", "$11-$25k", "$25-$50k", ">$50k")))

patient_data$dzgroup <- as.numeric(factor(patient_data$dzgroup, levels = c("ARF/MOSF w/Sepsis", "CHF", "COPD", "Cirrhosis", "Colon Cancer", "Coma", "Lung Cancer", "MOSF w/Malig")))

patient_data$dzclass <- as.numeric(factor(patient_data$dzclass, levels = c("ARF/MOSF", "COPD/CHF/Cirrhosis", "Cancer", "Coma")))

patient_data$ca <- as.numeric(factor(patient_data$ca, levels = c("no", "yes", "metastatic")))

patient_data$sfdm2 <- as.numeric(factor(patient_data$sfdm2, levels = c("no(M2 and SIP pres)", "adl>=4 (>=5 if sur)", "SIP>=30", "Coma or Intub","<2 mo. follow-up" )))

head(patient_data[, c("sex", "race", "income", "dzgroup", "dzclass", "ca", "sfdm2")])

patient_data
```

```{r}
summary(patient_data)
table(is.na(patient_data))  # Vérifier les valeurs manquantes

```

Travaille sur la variable target dead : 

```{r}
colonnes_a_supprimer <- c("sfdm2", "hospdead")
patient_data <- patient_data[, !(names(patient_data) %in% colonnes_a_supprimer)]
head(patient_data)
```
```{r}
library(corrplot)
cor_matrix <- cor(patient_data)
cor_matrix[abs(cor_matrix) < 0.6] <- 0
corrplot(cor_matrix, "color")
```
```{r}
library(dplyr) # Manipulation des données
library(caret) # Préparation des données et séparation en ensembles
library(pROC) # Analyse ROC et calcul de l'AUC
library(ggplot2) # Visualisation des résultats

```


```{r}
set.seed(123) 

trainIndex <- createDataPartition(patient_data$death, p = 0.8, list = FALSE)

train_data <- patient_data[trainIndex, ]

test_data <- patient_data[-trainIndex, ]
```

```{r}
res = glm(patient_data$death ~., data=patient_data, family="binomial")
mod_sum = summary(res)
mod_sum
```

